---
title: "Connor Nguyen"
output: html_notebook
---
# Homework 6
## conceptual
### excercise 1
#### a)
complete linkage

```{r}
a <- list()
a$merge <- matrix(c(-1, -2,
                    -3, -4,
                     1,  2), nc=2, byrow=TRUE ) 
a$height <- c(0.3, 0.5, 0.8)    # define merge heights
a$order <- 1:4              # order of leaves(trivial if hand-entered)
a$labels <- c(1,4,2,3)   # labels of leaves
class(a) <- "hclust"        # make it an hclust object
plot(a)                  
```
#### b)
single linkage
```{r}
a <- list()
a$merge <- matrix(c(-1, -2,
                    1, -3,
                    
                     2,  -4), nc=2, byrow=TRUE ) 
a$height <- c(0.3, 0.4, 0.45)    # define merge heights
a$order <- 1:4              # order of leaves(trivial if hand-entered)
a$labels <- c(1,4,2,3)   # labels of leaves
class(a) <- "hclust"        # make it an hclust object
plot(a)                  
```
#### c)
for complete linkage, the first cluster will include 1 and 4 and the second cluster will include 2,3

#### d)
for single linkage, the first cluster will include 1,4,2 and the second cluster will include 3

#### e)

```{r}
a <- list()
a$merge <- matrix(c(-3, -4,
                    -1, -2,
                     1,  2), nc=2, byrow=TRUE ) 
a$height <- c(0.3, 0.5, 0.8)    # define merge heights
a$order <- 1:4              # order of leaves(trivial if hand-entered)
a$labels <- c(2,3,1,4)   # labels of leaves
class(a) <- "hclust"        # make it an hclust object
plot(a)                  
```
### excercise 2:
#### a) 
single linkage

$$
\begin{matrix} 
 && 1 & 2 & 3 & 4 & {5,7}&6  &8 \\
\\
2&&0.66
\\
3&&1.08&1.70
\\
4&&0.97 &1.52&1.15
\\
5,7&&2.02&1.48&2.80&2.98
\\
6&& 0.36&0.96&1.00&0.61 &2.38
\\
8&&1.46&1.04&2.16&2.43&0.65&1.82
\\
9&&1.96 &2.01&1.91&2.77& 1.75&2.25&1.37

\end{matrix}

$$

#### b) 
average linkage

$$
\begin{matrix} 
 && 1 & 2 & 3 & 4 & {5,7}&6  &8 \\
\\
2&&0.66
\\
3&&1.08&1.70
\\
4&&0.97 &1.52&1.15
\\
5,7&& 2.13& 1.61&2.865 &3.09
\\
6&& 0.36&0.96&1.00&0.61 &2.485 
\\
8&&1.46&1.04&2.16&2.43&0.72&1.82
\\
9&&1.96 &2.01&1.91&2.77& 1.795&2.25&1.37

\end{matrix}

$$

### excercise 3
#### a)

there is not enough information to tell since the distance of 1 to 3 could be short and the distance between 1 to 2 is large.
compare those distance to the distance from 4 to 5.

#### b)

since there is only 2 data point, no matter what linkage you use, there is going to be the same height.

### excercise 4
### a)

```{r}
library('factoextra')
data(USArrests)
ex4.data=USArrests
hclust(dist(ex4.data), method= 'complete')
plot(hclust(dist(ex4.data), method= 'complete'))
```

#### b)

```{r}
cutree(hclust(dist(ex4.data), method= 'complete'), k=3)
```

#### c)

```{r}
ex4c.data=scale(USArrests)
hclust(dist(ex4c.data), method= 'complete')
plot(hclust(dist(ex4.data), method= 'complete'))
```

#### d)

```{r}
fviz_nbclust(ex4c.data, hcut,
             hc_method = "complete")
```
the optimal value of k is 2

#### e)

```{r}
cutree(hclust(dist(ex4c.data), method= 'complete'), k=2)
```

```{r}
ex4c.data[cutree(hclust(dist(ex4c.data), method= 'complete'), k=2)==1, ]

```

```{r}
ex4c.data[cutree(hclust(dist(ex4c.data), method= 'complete'), k=2)==2, ]
```

the 2 cluster are relatively far with respect to all variable since in cluster 1 all variable are positive and cluster 2 all variable are negative.

#### f)

there is a good reason to scale the data before hand because the variable could be large and the distance between cluster could change drastically

### Excercise 5
#### a)
 
 manually import the data
 
#### b)

```{r}
plot(hclust(dist(Ch10Ex11), method= 'complete'))
```
```{r}
plot(hclust(dist(Ch10Ex11), method= 'single'))
```
```{r}
plot(hclust(dist(Ch10Ex11), method= 'average'))
```

```{r}
fviz_nbclust(Ch10Ex11, hcut,
             hc_method = "complete")
```
for the complete linkage, the optimal K is 2

```{r}
fviz_nbclust(ex4c.data, hcut,
             hc_method = "single")
```
for single linkage, the optimal k is also 2

```{r}
fviz_nbclust(ex4c.data, hcut,
             hc_method = "average")
```
for average linkage, the optimal k is also 2

the result of optimal K does not depend on which linkage type I used















